---
output: 
  github_document: 
    toc: TRUE
---
   
<!-- README.md is generated from README.Rmd. Please edit that file --> 
 
   
```{r, include = FALSE}
knitr::opts_chunk$set(      
  collapse = TRUE,
  warning = FALSE,  
  comment = "#>",
  message = F
)        

```


```{r, results = 'asis', echo = F}
library(magrittr)
webpages <- fs::dir_ls(type = "file", 
                       recurse = T, 
                       glob = "*.csv|*.csv") %>% 
  rev()

file_name_wo_extension <- webpages %>% 
  stringr::str_remove(".html|.pdf")
 
webpages %>%    
  paste0("a. [",
         . ,
         "]",
         "(https://raw.githubusercontent.com/ggplot2-extenders/ggplot2-extension-scrapers/refs/heads/main/", ., 
         ")\n") %>% 
  cat()     
```

The repo contains code to characterize the ggplot2 extension ecosystem.  A couple of projects motivate this:

- [CRAN task views grammar of graphics (or ggplot2 extension)](https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82)
- [JSM 'Who are the ggplot2 extenders'](https://evamaerey.github.io/ggplot2-extension-ecosystem/)
 




 
 




# 1. `tools::CRAN_package_db` that are `^gg|^GG|gg$|GG$` w/ ggplot2 depend or import

Code/ideas: June, Joyce, Pepijn, Gina 

  
```{r}
df <- tools::CRAN_package_db() |> 
  dplyr::filter(
    stringr::str_detect(Package, "^gg|^GG|gg$|GG$"),
    stringr::str_detect(Depends, "ggplot2") | 
    stringr::str_detect(Imports, "ggplot2")
  ) 

df |> tibble::tibble()

df$Package

readr::write_csv(df, "gg_w_ggplot2_depends_or_imports_cran.csv")
```  


# 2. `yaml::read_yaml` and `httr2` to parse extension gallery `gallery _config.yml` file

Code/ideas: Pepijn, Joyce, Gina, (Probably others)



```{r}
df <-
  "https://raw.githubusercontent.com/ggplot2-exts/gallery/refs/heads/gh-pages/_config.yml" |>
  httr2::request() |>
  httr2::req_perform() |>
  httr2::resp_body_string() |>
  (\(x) yaml::read_yaml(text = x))() |>
  _$widgets |>
  dplyr::bind_rows()

df$name

df

readr::write_csv(df, "gg_extension_pkgs_gallery.csv")


```


# 3. `gh::gh` Keep in mind in terms: github contributors

"Re finding GitHub users from packages, the repo owner is not always the only/main contributor, especially if it's an org"

Code/ideas: Carl Suster @arcresu

https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82#discussioncomment-12469510

```{r}
gh_contributors <- function(repo) {
  resp <- gh::gh("GET /repos/{repo}/contributors", repo = repo)
  total_contributions <- sum(sapply(resp, \(x) x$contributions))
  resp <- Filter(\(x) x$type == "User", resp) # exclude bots
  resp <- Filter(\(x) x$contributions > total_contributions/5, resp) # with at least 20% of contributions
  sapply(resp, \(x) stringr::str_trim(x$login))
}

gh_contributors("cidm-ph/ggmapinset")
gh_contributors("YuLab-SMU/ggfun")

```


# 4. `universe::global_search` with exported function pattern identification

Teun  https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82#discussioncomment-12479880



```{r, eval = F}
# install.packages("universe", repos = "https://ropensci.r-universe.dev")

# I'm aware there should be ~7k/8k packages with ggplot2 as dependency.
packages <- universe::global_search(query = 'needs:ggplot2', limit = 10000L)
out_file <- "universe_ggplot2_depends_function_exports.csv"

# Ensure I have a 'data' folder with the file I'll need
if (!fs::file_exists(out_file)) {
	dir <- fs::path_dir(out_file)
	if (!fs::dir_exists(dir)) {
		fs::dir_create(dir)
	}
	fs::file_create(out_file)
}

# Read current data if it is cached
current_data <- data.table::fread(out_file)

data <- lapply(packages$results, function(result) {

	name <- result$Package
	universe <- result$`_user`

	# We're going to skip this package if we've already seen it. Potentially,
	# we'd be skipping packages with duplicate names, but that shouldn't occur
	# too often.
	if (name %in% current_data$name) {
		return()
	}

	# The information we want, the exported functions, is not available in
	# the results we already have. We need to a package specific query
	# to get the `_exports` field
	details <- universe::universe_one_package(universe, package = name)
	exports <- unlist(details$`_exports`) %||% NA_character_

	# Format as data.frame
	df <- data.frame(
		name = name,
		universe = universe,
		export = exports
	)

	# Write to file directly. Combined with the skip mechanism above, we're
	# effectively caching every result
	data.table::fwrite(df, out_file, append = TRUE)
})


library(dplyr)
library(ggplot2)
library(scales)

file <- "universe_ggplot2_depends_function_exports.csv"

data <- data.table::fread(file) |>
	filter(nzchar(export)) |>
	filter(!startsWith(name, "RcmdrPlugin")) |>
	mutate(class = case_when(
		startsWith(export, "geom_")     ~ "geom",
		startsWith(export, "stat_")     ~ "stat",
		startsWith(export, "scale_")    ~ "scale",
		startsWith(export, "coord_")    ~ "coord",
		startsWith(export, "facet_")    ~ "facet",
		startsWith(export, "guide_")    ~ "guide",
		startsWith(export, "position_") ~ "position",
		startsWith(export, "draw_key_") ~ "key",
		startsWith(export, "element_")  ~ "element",
		startsWith(export, "theme_")    ~ "theme",
		startsWith(export, "Geom")      ~ "Geom class",
		startsWith(export, "Stat")      ~ "Stat class",
		startsWith(export, "Scale")     ~ "Scale class",
		startsWith(export, "Coord")     ~ "Coord class",
		startsWith(export, "Facet")     ~ "Facet class",
		startsWith(export, "Guide")     ~ "Guide class",
		startsWith(export, "Position")  ~ "Position class",
		.default = ""
	)) |>
	mutate(pattern = case_when(
		startsWith(name, "gg")   ~ "gg-prefix",
		startsWith(name, "tidy") ~ "tidy-prefix",
		endsWith(name, "themes") ~ "themes-suffix",
		.default = ""
	))

write.csv(gg_pkgs_data.df, file = "gg-pkgs-data.csv")

```



# 5. `pkgdiff` to look at patterns also?

Pedro Aphalo

https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82#discussioncomment-12582326

```{r, eval = F}
library(pkgdiff)
library(lubridate)
library(dplyr)



# rm(list = ls(pattern = "*"))

pkg_stability_row <- function(pkg, releases = NULL, months = NULL) {
  temp <- pkg_stability(pkg = pkg, releases = releases, months = months)
  temp.df <- as.data.frame(temp[c(1:7)])
  temp.df$num.funs <- temp$StabilityData$TF[1L]
  temp.df$Size <- temp$StabilityData$Size[1L]
  temp.df
}

pkg_gg_functions <- function(pkg) {
  gg.funs <- list(PackageName = pkg)
  temp <- pkg_info(pkg = pkg, ver = "latest") # latest in CRAN! (ignores local)
  fun.names <- unique(names(temp$Functions))
  gg.funs$num.geoms <- sum(grepl("^geom_", fun.names))
  gg.funs$num.stats <- sum(grepl("^stat_", fun.names))
  gg.funs$num.scales <- sum(grepl("^scale_", fun.names))
  gg.funs$num.positions <- sum(grepl("^position_", fun.names))
  gg.funs$num.coords <- sum(grepl("^coord_", fun.names))
  gg.funs$num.drawkeys <- sum(grepl("^draw_key_", fun.names))
  gg.funs$num.guides <- sum(grepl("^guide_", fun.names))
  gg.funs$num.labellers <- sum(grepl("^label_", fun.names))
  gg.funs$num.themes <- sum(grepl("^theme_", fun.names))
  gg.funs$num.theme.elements <- sum(grepl("^element_", fun.names))
  gg.funs$num.ggplots <- sum(grepl("^ggplot", fun.names))
  gg.funs$num.autoplots <- sum(grepl("^autoplot", fun.names))
  gg.funs$num.autolayers <- sum(grepl("^autolayer", fun.names))
  as.data.frame(gg.funs)
}

csv <- "https://raw.githubusercontent.com/ggplot2-extenders/ggplot2-extensions-cran-task-view/refs/heads/main/gg-pkgs.csv"

# downloaded previously from GitHub
gg_pkgs_list.df <- read.csv(csv)
gg_packages <- gg_pkgs_list.df$Package
length(gg_packages)

# 'pkgdiff' gets package data from CRAN
cran_pkgs <- available.packages(repos = c(CRAN = "https://cran.rstudio.com/"))
cran_pkgs <- cran_pkgs[ , "Package"]

gg_packages <- intersect(gg_packages, cran_pkgs)
length(gg_packages)

## run only if cached
# chached_packages <- pkg_cache()
# gg_packages <- intersect(na.omit(chached_packages$Package), gg_packages)
# length(gg_packages)

# even fewer packages for testing
# gg_packages <- gg_packages[1:5]

# all work lost if functions fail to return a value
# stability.ls <- lapply(gg_packages, pkg_stability_row)
# stability.df <- bind_rows(stability.ls)
# 
# functions.ls <- lapply(gg_packages, pkg_gg_functions)
# functions.df <- bind_rows(functions.ls)

# use a for loop instead so that results are not all lost when the function errors.
# It can take quite a long time to run.

if (!exists("stability.ls")) stability.ls <- list()
if (!exists("functions.ls")) functions.ls <- list()
pkgs_done <- intersect(names(stability.ls), names(functions.ls))
pkgs_to_do <- setdiff(gg_packages, pkgs_done)
length(pkgs_to_do)

for (pkg in pkgs_to_do) {
  temp1 <- pkg_stability_row(pkg = pkg)
  if (nrow(temp1)) {
    stability.ls[[pkg]] <- temp1
  }
  temp2 <- pkg_gg_functions(pkg = pkg)
  if (nrow(temp2)) {
    functions.ls[[pkg]] <- temp2
  }
}

stability.df <- bind_rows(stability.ls)
functions.df <- bind_rows(functions.ls)

stability.df |> tibble::tibble()
functions.df |> tibble::tibble()

gg_pkgs_data.df <- full_join(stability.df, functions.df)


```


