---
output: 
  github_document: 
    toc: TRUE
---
   
<!-- README.md is generated from README.Rmd. Please edit that file --> 
 
   
```{r, include = FALSE}
knitr::opts_chunk$set(      
  collapse = TRUE,
  warning = FALSE,  
  comment = "#>",
  message = F
)        

```


```{r, results = 'asis', echo = F}
library(magrittr)
webpages <- fs::dir_ls(type = "file", 
                       recurse = T, 
                       glob = "*.csv|*.csv") %>% 
  rev()

file_name_wo_extension <- webpages %>% 
  stringr::str_remove(".html|.pdf")
 
webpages %>%    
  paste0("a. [",
         . ,
         "]",
         "(https://raw.githubusercontent.com/ggplot2-extenders/ggplot2-extension-scrapers/refs/heads/main/", ., 
         ")\n") %>% 
  cat()     
```

The repo contains code to characterize the ggplot2 extension ecosystem.  A couple of projects motivate this:

- [CRAN task views grammar of graphics (or ggplot2 extension)](https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82)
- [JSM 'Who are the ggplot2 extenders?'](https://evamaerey.github.io/ggplot2-extension-ecosystem/)
 

```{r, eval = F}
devtools::create(".")
```







# 2. `yaml::read_yaml` and `httr2` to parse extension gallery `gallery _config.yml` file 148 packages

Code/ideas: Pepijn, Joyce, Gina, (Probably others)

```{r}
gg_gallery_pkgs <-
  "https://raw.githubusercontent.com/ggplot2-exts/gallery/refs/heads/gh-pages/_config.yml" |>
  httr2::request() |>
  httr2::req_perform() |>
  httr2::resp_body_string() |>
  (\(x) yaml::read_yaml(text = x))() |>
  _$widgets |>
  dplyr::bind_rows() |>
  dplyr::rename(package = name)

gg_gallery_pkgs |> tibble::tibble()
gg_gallery_pkgs$package

usethis::use_data(gg_gallery_pkgs, overwrite = T)

```

# 1. `tools::CRAN_package_db` that are `^gg|^GG|gg$|GG$` or found in gallery w/ ggplot2 depend or import (242 packages)

Code/ideas: June, Joyce, Pepijn, Gina 

  
```{r}
cran_gg_w_ggplot2_depends_or_imports <- tools::CRAN_package_db() |> 
  dplyr::filter(
    stringr::str_detect(Package, "^gg|^GG|gg$|GG$") | 
      Package %in% gg_gallery_pkgs$package,
    stringr::str_detect(Depends, "ggplot2") | 
      stringr::str_detect(Imports, "ggplot2")
  ) |>
  dplyr::mutate(ind_gallery = Package %in% gg_gallery_pkgs$package) |>
  janitor::clean_names()

cran_gg_w_ggplot2_depends_or_imports |> tibble::tibble() 
names(cran_gg_w_ggplot2_depends_or_imports) 
cran_gg_w_ggplot2_depends_or_imports$package |> sort()

usethis::use_data(cran_gg_w_ggplot2_depends_or_imports, overwrite = T)



```  

```{r}
# devtools::check()
# devtools::install(".", upgrade = "never")
```

```{r}
knitr::opts_chunk$set(eval= F)

```


```{r, fig.width=15, fig.height=15, eval = F, include=F}
library(tidyverse)
library(ggplot2.extension.scrapers)
library(ggraph)

df <- cran_gg_w_ggplot2_depends_or_imports

cran_gg_w_ggplot2_depends_or_imports |> 
  mutate(author = str_split(author, pattern = "\\n")) |>
  unnest(author) |>
  select(author, package) |>
  mutate(author = str_remove(author, "\\[.+")) |>
  mutate(author = str_trim(author)) |>
  filter(!(author |> str_detect("orcid"))) ->
package_authors_edge_list

package_authors_edge_list |>
  tidygraph::as_tbl_graph() |>
  ggraph(layout = "kk") +
  geom_node_text(aes(label = name, color = name %in% cran_gg_w_ggplot2_depends_or_imports$package)) + 
  geom_edge_link(alpha = .2)
  
library(ggram)

package_authors_edge_list |>
  # ungroup() |> 
  mutate(num_authorships = n(), .by = author) |>
  filter(num_authorships >= 4) |>
  tidygraph::as_tbl_graph() |>
  ggraph(layout = "fr") +
  geom_edge_link(alpha = .2) + 
  geom_node_label(aes(label = name, 
                      fill = name %in% c("Hadley Wickham", "Jonathan Sidi", 
                                          "Jonathan Carroll", "Teun van den Brand", "Claus O. Wilke",
                                          "David Sjoberg", "Malcolm Barrett", "Di Cook", 
                                         "Elio Campitelli", "Joseph Larmarange",
                                         "David Kahle", "Winston Chang", "Thomas Lin Pedersen", "Matthew Kay"),
                  color = name %in% package_authors_edge_list$author)) + 
  theme(legend.position = "none") + 
  scale_fill_manual(values = c("grey90", "white"))
    
```


# 3. `gh::gh` Keep in mind in terms: github contributors

"Re finding GitHub users from packages, the repo owner is not always the only/main contributor, especially if it's an org"

Code/ideas: Carl Suster @arcresu

https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82#discussioncomment-12469510

```{r}
gh_contributors <- function(repo = "cidm-ph/ggmapinset") {
  resp <- gh::gh("GET /repos/{repo}/contributors", repo = repo)
  total_contributions <- sum(sapply(resp, \(x) x$contributions))
  resp <- Filter(\(x) x$type == "User", resp) # exclude bots
  resp <- Filter(\(x) x$contributions > total_contributions/5, resp) # with at least 20% of contributions
  sapply(resp, \(x) stringr::str_trim(x$login))
}

gh_contributors("cidm-ph/ggmapinset")
gh_contributors("YuLab-SMU/ggfun")
gh_contributors("tidyverse/ggplot2")
gh_contributors("AllanCameron/geomtextpath")

```


# 4. `universe::global_search` with exported function pattern identification

Teun  https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82#discussioncomment-12479880



```{r, eval = F}
# install.packages("universe", repos = "https://ropensci.r-universe.dev")

# I'm aware there should be ~7k/8k packages with ggplot2 as dependency.
packages <- universe::global_search(query = 'needs:ggplot2', limit = 10000L)
out_file <- "universe_ggplot2_depends_function_exports.csv"

# Ensure I have a 'data' folder with the file I'll need
if (!fs::file_exists(out_file)) {
	dir <- fs::path_dir(out_file)
	if (!fs::dir_exists(dir)) {
		fs::dir_create(dir)
	}
	fs::file_create(out_file)
}

# Read current data if it is cached
current_data <- data.table::fread(out_file)

data <- lapply(packages$results, function(result) {

	name <- result$Package
	universe <- result$`_user`

	# We're going to skip this package if we've already seen it. Potentially,
	# we'd be skipping packages with duplicate names, but that shouldn't occur
	# too often.
	if (name %in% current_data$name) {
		return()
	}

	# The information we want, the exported functions, is not available in
	# the results we already have. We need to a package specific query
	# to get the `_exports` field
	details <- universe::universe_one_package(universe, package = name)
	exports <- unlist(details$`_exports`) %||% NA_character_

	# Format as data.frame
	df <- data.frame(
		name = name,
		universe = universe,
		export = exports
	)

	# Write to file directly. Combined with the skip mechanism above, we're
	# effectively caching every result
	data.table::fwrite(df, out_file, append = TRUE)
})


library(dplyr)
library(ggplot2)
library(scales)

file <- "universe_ggplot2_depends_function_exports.csv"

data <- data.table::fread(file) |>
	filter(nzchar(export)) |>
	filter(!startsWith(name, "RcmdrPlugin")) |>
	mutate(class = case_when(
		startsWith(export, "geom_")     ~ "geom",
		startsWith(export, "stat_")     ~ "stat",
		startsWith(export, "scale_")    ~ "scale",
		startsWith(export, "coord_")    ~ "coord",
		startsWith(export, "facet_")    ~ "facet",
		startsWith(export, "guide_")    ~ "guide",
		startsWith(export, "position_") ~ "position",
		startsWith(export, "draw_key_") ~ "key",
		startsWith(export, "element_")  ~ "element",
		startsWith(export, "theme_")    ~ "theme",
		startsWith(export, "Geom")      ~ "Geom class",
		startsWith(export, "Stat")      ~ "Stat class",
		startsWith(export, "Scale")     ~ "Scale class",
		startsWith(export, "Coord")     ~ "Coord class",
		startsWith(export, "Facet")     ~ "Facet class",
		startsWith(export, "Guide")     ~ "Guide class",
		startsWith(export, "Position")  ~ "Position class",
		.default = ""
	)) |>
	mutate(pattern = case_when(
		startsWith(name, "gg")   ~ "gg-prefix",
		startsWith(name, "tidy") ~ "tidy-prefix",
		endsWith(name, "themes") ~ "themes-suffix",
		.default = ""
	))


universe_gg_depends_function_exports <- readr::read_csv("universe_ggplot2_depends_function_exports.csv")

usethis::use_data()

```



# 5. `pkgdiff` to look at patterns also?

Pedro Aphalo

https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/82#discussioncomment-12582326

```{r, eval = F}
library(pkgdiff)
library(lubridate)
library(dplyr)



# rm(list = ls(pattern = "*"))

pkg_stability_row <- function(pkg, releases = NULL, months = NULL) {
  temp <- pkg_stability(pkg = pkg, releases = releases, months = months)
  temp.df <- as.data.frame(temp[c(1:7)])
  temp.df$num.funs <- temp$StabilityData$TF[1L]
  temp.df$Size <- temp$StabilityData$Size[1L]
  temp.df
}

pkg_gg_functions <- function(pkg) {
  gg.funs <- list(PackageName = pkg)
  temp <- pkg_info(pkg = pkg, ver = "latest") # latest in CRAN! (ignores local)
  fun.names <- unique(names(temp$Functions))
  gg.funs$num.geoms <- sum(grepl("^geom_", fun.names))
  gg.funs$num.stats <- sum(grepl("^stat_", fun.names))
  gg.funs$num.scales <- sum(grepl("^scale_", fun.names))
  gg.funs$num.positions <- sum(grepl("^position_", fun.names))
  gg.funs$num.coords <- sum(grepl("^coord_", fun.names))
  gg.funs$num.drawkeys <- sum(grepl("^draw_key_", fun.names))
  gg.funs$num.guides <- sum(grepl("^guide_", fun.names))
  gg.funs$num.labellers <- sum(grepl("^label_", fun.names))
  gg.funs$num.themes <- sum(grepl("^theme_", fun.names))
  gg.funs$num.theme.elements <- sum(grepl("^element_", fun.names))
  gg.funs$num.ggplots <- sum(grepl("^ggplot", fun.names))
  gg.funs$num.autoplots <- sum(grepl("^autoplot", fun.names))
  gg.funs$num.autolayers <- sum(grepl("^autolayer", fun.names))
  as.data.frame(gg.funs)
}

csv <- "https://raw.githubusercontent.com/ggplot2-extenders/ggplot2-extensions-cran-task-view/refs/heads/main/gg-pkgs.csv"

# downloaded previously from GitHub
gg_pkgs_list.df <- read.csv(csv)
gg_packages <- gg_pkgs_list.df$Package
length(gg_packages)

# 'pkgdiff' gets package data from CRAN
cran_pkgs <- available.packages(repos = c(CRAN = "https://cran.rstudio.com/"))
cran_pkgs <- cran_pkgs[ , "Package"]

gg_packages <- intersect(gg_packages, cran_pkgs)
length(gg_packages)

## run only if cached
# chached_packages <- pkg_cache()
# gg_packages <- intersect(na.omit(chached_packages$Package), gg_packages)
# length(gg_packages)

# even fewer packages for testing
# gg_packages <- gg_packages[1:5]

# all work lost if functions fail to return a value
# stability.ls <- lapply(gg_packages, pkg_stability_row)
# stability.df <- bind_rows(stability.ls)
# 
# functions.ls <- lapply(gg_packages, pkg_gg_functions)
# functions.df <- bind_rows(functions.ls)

# use a for loop instead so that results are not all lost when the function errors.
# It can take quite a long time to run.

if (!exists("stability.ls")) stability.ls <- list()
if (!exists("functions.ls")) functions.ls <- list()
pkgs_done <- intersect(names(stability.ls), names(functions.ls))
pkgs_to_do <- setdiff(gg_packages, pkgs_done)
length(pkgs_to_do)

for (pkg in pkgs_to_do) {
  temp1 <- pkg_stability_row(pkg = pkg)
  if (nrow(temp1)) {
    stability.ls[[pkg]] <- temp1
  }
  temp2 <- pkg_gg_functions(pkg = pkg)
  if (nrow(temp2)) {
    functions.ls[[pkg]] <- temp2
  }
}

stability.df <- bind_rows(stability.ls)
functions.df <- bind_rows(functions.ls)

stability.df |> tibble::tibble()
functions.df |> tibble::tibble()

gg_pkgs_data.df <- full_join(stability.df, functions.df)


```


